---
title: 'EVRN 745 Assignment 3: gganimate guide'
author: "Christopher Wheeler"
date: '2022-10-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1) Goals of the package
The main goal of gganimate is to extend to ggplot syntax and aesthetic scheme to the generation of animated graphs. It does this by adding a few new types of functions dealing with implemeting different types of animations and rendering the resultant `gganim` objects as GIF files. These functions are appended to the ggplot call using the expected ` + ` operator. Here are some examples of these functions and their uses:

`transition_*()` defines how the data should be spread out and how it relates to itself across time.
`view_*() defines `how the positional scales should change along the animation.
`shadow_*() defines` how data from other points in time should be presented in the given point in time.
`enter_*()/exit_*()` defines how new data should appear and how old data should disappear during the course of the animation.
`ease_aes() `defines how different aesthetics should be eased during transitions.


## 2) Examples demonstrating the functionality of the package

```{r}
library(tidyverse)
library(gganimate)
library(gifski)

ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  # Here comes the gganimate code
  transition_states(
    gear,
    transition_length = 2,
    state_length = 1) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('sine-in-out') + 
   theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size = 1)) + 
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
```



## 3) An example figure

```{r}
library(gapminder)

ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear') + 
  theme_bw() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size = 1)) + 
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
```

